### 2015-07-13 narrative/data-reduction.tex
Based on title its SIRIUS not SIRUS.

The classical workflow where the entire dataset is written to storage for
later analysis will no longer be viable in exascale, simply because the amount of generated data will 
be too large due to capacity and performance reasons.
In the future, it will be vital to take advantage of \textit{a priori}
information when 1) writing and reading from the SSIO layer in order to gain
higher performance and predictability, and to prioritize the data that
is most useful for end users so that I/O can be finished in the time available, and 2)
performing {\em in situ} operations and analysis before storing the information.
In the fusion use case described earlier which shares commonalities with many other
DOE applications, \textit{a priori} information can indeed be provided by application
scientists regarding which data should be sent over to the SSIO layer, which we refer
to in our proposal as {\bf SIRUS}, so that
minimally, the most science relevant data can be available for subsequent analysis.
This allows the science to be done even when the storage is in busy state - we believe
in a shared multi-user environment, it occurs often \cite{liu_hotstorage}, and gives users
high performance variability.
In this project, this capability of prioritizing data is accomplished by data refactoring, 
which includes re-organization and reductions of the data.
There are many methodologies to refactor data and the best choice will generally
be application dependent. However, our observation is that, once the choice is settled for an application, 
it will not typically change from run to run.

@@

Deep application knowledge means one can sometimes achieve drastically superior
reduction in the total data compared with what one might achieve otherwise.
However, even in the absence of such high level knowledge, we must equip SIRUS with generic data reduction techniques and re-organization
techniques. For instance, certain basic information on the semantics of the
data is needed and must be supported by the overall infrastructure. Likewise,
information regarding typical access patterns for data in a given application
should be placed in SIRUS. Scientific applications often read and
analyze data in a structured fashion and failure to respect this pattern, or to
exploit it through prefetching, can significantly increase the overall
execution time of the application. Accurate prefetching can only be achieved
with access pattern analysis (\S\ref{sec:managing-data-life}).

@@

To implement these plug-in capabilities, we will build upon our existing data staging 
capability and further allow data to be
efficiently staged, re-organized and mapped to a multi-tiered storage. Various
processing scenarios such as in-situ, in-transit or a combination of the two
will be researched and developed to meet our needs.
%The open questions that will be addressed include: 
%1) do we move the data over to
%other nodes to then reduce and re-organize, 2) do we first reduce the data and
%then later re-organize this data on different nodes?, or 3) do we use different
%cores on a node to do data-refactoring.  
Similarly we will examine the cost of reading the data. In the case where only high importance being retrieved,
overall time to move the data from SIRUS to the
application is expected to fast. A key challenge is in the assessment of retrieving
both high and low important data that reside across storage tiers and possibly migrate to
new locations as a result of capacity pressure. There is a
clear penalty of having to discovery data (\S\ref{sec:naming-discovery}), take data in a hierarchical layout and then re-compose
this back to a single-layer view.  
%Additional research questions which must be
%addressed: Will some types of data (e.g. PIC data) be more
%conducive  for data-refactoring than others? How much semantic-domain knowledge
%will we need to get a dramatic improvement of the storage system over not doing
%anything special?

@@

\paragraph{Challenges:} Maximal data reduction in SIRUS
would result from storing only the code and the input parameters, and simply
recalculating all of the results. This extreme is not feasible in practice
owing to the size of the computation involved, but sheds light on the path to
achieving a step change in data reduction. We shall investigate the possibility
of writing a minimal set of information, and placing an ``auditor'' code in a
container-like environment so that data can be re-generated with a {\em local}
approximation of the full simulation. This is an entirely new approach to
tackle the data overload by using preprocessors in which information is not
simply stored in the form of raw data, but rather in the form of appropriate
parameterized algorithms. An ``auditor'' is a code which mimics the exascale simulation, 
but typically can run using less degrees of freedom. This means that the code is a good
local (in space and time) approximation of the exascale simulation and will be much less
expensive to execute (so that it can be run in situ), and follows the behavior of the simulation
for short periods of time. This allows us to ``audit'' the simulation and when the differences 
become large in a local region of space, we write out the results of the data in this local region.
The hope is that this will be done only locally and that with a good ``auditor''  this will not happen
too frequently. This then allows us to use the ``auditor'' to reproduce results during the analysis 
process using an inexpensive computation and put less pressure on SIRUS
%
% 
%   A priori information on the nature of the data is
%used to choose the basic algorithm while the actual values of the data are used
%to tune the parameters used in the reconstruction procedures. We shall refer to
%this paradigm as an Auditor.  We will examine the use of several auditors for
%the simulations which we will work with. One approach is to use a fluid-code to
%audit a kinetic code, which can be a good approximation over short periods of
%time.

>>>>

The classical workflow where the entire dataset is written to storage for
later analysis will no longer be viable in exascale, simply because the amount of generated data will 
be too large due to capacity and performance reasons.
In the future, it will be vital to take advantage of \textit{a priori}
information when 1) writing and reading from the SSIO layer in order to gain
higher performance and predictability, and to prioritize the data that
is most useful for end users so that I/O can be finished in the time available, and 2)
performing {\em in situ} operations and analysis before storing the information.
In the fusion use case described earlier which shares commonalities with many other
DOE applications, \textit{a priori} information can indeed be provided by application
scientists regarding which data should be sent over to the SSIO layer, which we refer
to in our proposal as {\bf SIRIUS}, so that
minimally, the most science relevant data can be available for subsequent analysis.
This allows the science to be done even when the storage is in busy state - we believe
in a shared multi-user environment, it occurs often \cite{liu_hotstorage}, and gives users
high performance variability.
In this project, this capability of prioritizing data is accomplished by data refactoring, 
which includes re-organization and reductions of the data.
There are many methodologies to refactor data and the best choice will generally
be application dependent. However, our observation is that, once the choice is settled for an application, 
it will not typically change from run to run.

@@

Deep application knowledge means one can sometimes achieve drastically superior
reduction in the total data compared with what one might achieve otherwise.
However, even in the absence of such high level knowledge, we must equip SIRIUS with generic data reduction techniques and re-organization
techniques. For instance, certain basic information on the semantics of the
data is needed and must be supported by the overall infrastructure. Likewise,
information regarding typical access patterns for data in a given application
should be placed in SIRIUS. Scientific applications often read and
analyze data in a structured fashion and failure to respect this pattern, or to
exploit it through prefetching, can significantly increase the overall
execution time of the application. Accurate prefetching can only be achieved
with access pattern analysis (\S\ref{sec:managing-data-life}).

@@

To implement these plug-in capabilities, we will build upon our existing data staging 
capability and further allow data to be
efficiently staged, re-organized and mapped to a multi-tiered storage. Various
processing scenarios such as in-situ, in-transit or a combination of the two
will be researched and developed to meet our needs.
%The open questions that will be addressed include: 
%1) do we move the data over to
%other nodes to then reduce and re-organize, 2) do we first reduce the data and
%then later re-organize this data on different nodes?, or 3) do we use different
%cores on a node to do data-refactoring.  
Similarly we will examine the cost of reading the data. In the case where only high importance being retrieved,
overall time to move the data from SIRIUS to the
application is expected to fast. A key challenge is in the assessment of retrieving
both high and low important data that reside across storage tiers and possibly migrate to
new locations as a result of capacity pressure. There is a
clear penalty of having to discovery data (\S\ref{sec:naming-discovery}), take data in a hierarchical layout and then re-compose
this back to a single-layer view.  
%Additional research questions which must be
%addressed: Will some types of data (e.g. PIC data) be more
%conducive  for data-refactoring than others? How much semantic-domain knowledge
%will we need to get a dramatic improvement of the storage system over not doing
%anything special?

@@

\paragraph{Challenges:} Maximal data reduction in SIRIUS
would result from storing only the code and the input parameters, and simply
recalculating all of the results. This extreme is not feasible in practice
owing to the size of the computation involved, but sheds light on the path to
achieving a step change in data reduction. We shall investigate the possibility
of writing a minimal set of information, and placing an ``auditor'' code in a
container-like environment so that data can be re-generated with a {\em local}
approximation of the full simulation. This is an entirely new approach to
tackle the data overload by using preprocessors in which information is not
simply stored in the form of raw data, but rather in the form of appropriate
parameterized algorithms. An ``auditor'' is a code which mimics the exascale simulation, 
but typically can run using less degrees of freedom. This means that the code is a good
local (in space and time) approximation of the exascale simulation and will be much less
expensive to execute (so that it can be run in situ), and follows the behavior of the simulation
for short periods of time. This allows us to ``audit'' the simulation and when the differences 
become large in a local region of space, we write out the results of the data in this local region.
The hope is that this will be done only locally and that with a good ``auditor''  this will not happen
too frequently. This then allows us to use the ``auditor'' to reproduce results during the analysis 
process using an inexpensive computation and put less pressure on SIRIUS
%
% 
%   A priori information on the nature of the data is
%used to choose the basic algorithm while the actual values of the data are used
%to tune the parameters used in the reconstruction procedures. We shall refer to
%this paradigm as an Auditor.  We will examine the use of several auditors for
%the simulations which we will work with. One approach is to use a fluid-code to
%audit a kinetic code, which can be a good approximation over short periods of
%time.

### 2015-07-13 narrative/intro.tex
Based on title its SIRIUS not SIRUS.

%%% Manish: I am not sure I fully understand the first part of the paragraph
%%% below.
The proposed work explores how application-specific information about the
importance of data will be used in conjunction with system and
resource-specific information to optimize the scientists' experience in
interacting with, and extracting knowledge and insight from their data.
%
%The overarching idea of this proposal is that data from applications can be characterized by different
%%levels of importance and different 
%levels of importance over time, which then can be used to guide the
% analysis and visualization later to access the most important data and
% deliver some level of knowledge without waiting for access to the whole
% original dataset. Information about the importance of the data at the
% application level, information about resource status and performance of
% the multi-level storage system, and flexible data reorganization and
% reduction operations at all levels can be utilized to provide a quality of
% service and allow users to gain insight in a timely manner.
The project combines knowledge and expertise 
from a wide range of sources including many leadership class applications using the Adaptable Input/Output System (ADIOS) framework,
the Sandia National Laboratory (SNL) Lightweight File Systems (LWFS) project and storage system knowledge, the metadata
management and storage systems knowledge at UCSC, and the middleware expertise
at Rutgers. We will integrate this diverse expertise to investigate how to create
a data, workflow, and application aware storage and I/O system, which we refer to as
{\bf SIRUS}.
%This project brings together a team with strong expertise in I/O middleware (ORNL, Rutgers),
%file system (SNL, UCSC) and storage (UCSC), and connects and
%coordinates these key storage components in a seamless fashion.
%
%\section{Scalable Storage Software Infrastructure}
%Our focus in this proposal is in Theme 2: {\em Scalable Storage Software Infrastructure} 
%We will focus on a state of the art SSIO solution to solve problems which researchers are 
%beginning to face and will make science on exascale systems
%problematic without new insights and solutions.
Our proposal will be responsive to Theme 2: {\em Scalable Storage Software Infrastructure} of the call, as summarized in Table~\ref{tab:call}.

>>>>

%%% Manish: I am not sure I fully understand the first part of the paragraph
%%% below.
The proposed work explores how application-specific information about the
importance of data will be used in conjunction with system and
resource-specific information to optimize the scientists' experience in
interacting with, and extracting knowledge and insight from their data.
%
%The overarching idea of this proposal is that data from applications can be characterized by different
%%levels of importance and different 
%levels of importance over time, which then can be used to guide the
% analysis and visualization later to access the most important data and
% deliver some level of knowledge without waiting for access to the whole
% original dataset. Information about the importance of the data at the
% application level, information about resource status and performance of
% the multi-level storage system, and flexible data reorganization and
% reduction operations at all levels can be utilized to provide a quality of
% service and allow users to gain insight in a timely manner.
The project combines knowledge and expertise 
from a wide range of sources including many leadership class applications using the Adaptable Input/Output System (ADIOS) framework,
the Sandia National Laboratory (SNL) Lightweight File Systems (LWFS) project and storage system knowledge, the metadata
management and storage systems knowledge at UCSC, and the middleware expertise
at Rutgers. We will integrate this diverse expertise to investigate how to create
a data, workflow, and application aware storage and I/O system, which we refer to as
{\bf SIRIUS}.
%This project brings together a team with strong expertise in I/O middleware (ORNL, Rutgers),
%file system (SNL, UCSC) and storage (UCSC), and connects and
%coordinates these key storage components in a seamless fashion.
%
%\section{Scalable Storage Software Infrastructure}
%Our focus in this proposal is in Theme 2: {\em Scalable Storage Software Infrastructure} 
%We will focus on a state of the art SSIO solution to solve problems which researchers are 
%beginning to face and will make science on exascale systems
%problematic without new insights and solutions.
Our proposal will be responsive to Theme 2: {\em Scalable Storage Software Infrastructure} of the call, as summarized in Table~\ref{tab:call}.
